Space-time adaptivity for parabolic evolution equations

### 2020-05-26 Latex/paper.tex
try try

This error estimator will drive our adaptive loop. Even without Assumption~\ref{assum:X-stable-splitting} however, we derive the following efficient and reliable \emph{global estimator}.
\begin{prop}
  \label{prop:global-apost}
  Let $\Xi^{\udelta}$ be a basis for $Y^{\udelta}$. Then, modulo some data-oscillation
  term we expect to be of higher order, for any $v^\delta \in X^\delta$,
  \[
    \|u - v^\delta\|_X^2 \eqsim \la (\mathbf A_s^{\udelta})^{-1}{\bf r}^\udelta_Y, {\bf r}^\udelta_Y \ra + \|u_0 - \gamma_0 v^\delta\|_H^2,
  \] 
  where $\mathbf A_s^{\udelta} := (A_s E_Y^{\udelta} \Xi^{\udelta})(E_Y^{\udelta} \Xi^{\udelta})$ and ${\bf r}^\udelta_Y := (g - B E_X^{\delta} v^\delta)(\Xi^\udelta)$.
  \jw{zouden we in staat zijn om de constantes hidden in $\eqsim$ hierboven, precies te maken? Mss de data oscillation erin?}
\end{prop}
\begin{proof}
  Define $e^\delta := u - E_X^\delta v^\delta$. From~\eqref{eqn:B-lis}, we derive
  \[
    \|e^\delta\|_X^2 \eqsim \|B e^\delta\|_{Y' \times H}^2
    = \|B e^\delta\|_{Y'}^2 + \|\gamma_0 e^\delta\|_H^2 = \|B e^\delta\|_{Y'}^2 + \|u_0 - \gamma_0 v^\delta\|_H^2,
  \]
  and using~\cite[Lemma 3.4]{Stevenson2020a} together with~\eqref{eqn:infinfsup}, we find
  \begin{align*}
    \|B e^\delta\|_{Y'}^2 &= \|e^\delta\|_Y^2 + \|\partial_t e^\delta\|_{Y'}^2 + \|\gamma_T e^\delta\|_H^2 - \|\gamma_0 e^\delta\|_H^2 \\
    &\eqsim \|e^\delta\|_Y^2 + \|\partial_t e^\delta\|_{{Y^\udelta}'}^2 + \|\gamma_T e^\delta\|_H^2 - \|\gamma_0 e^\delta\|_H^2 = \|B e^\delta\|_{{Y^\udelta}'}^2.
  \end{align*}
  The result then follows from
  \begin{align*}
    \|B e^\delta\|_{{Y^\udelta}'}^2 &= \|g - BE_X^\delta v^\delta\|_{{Y^\udelta}'}^2
    = \sup_{\mathclap{0 \not= w \in Y^\udelta}} \frac{[(g - Bv^\delta)(w)]^2}{\|w\|_Y^2} = ~~\sup_{\mathclap{0 \not= {\bf w} \in \R^{\# Y^\udelta}}} ~~\frac{[(g - Bv^\delta)({\bf w}^\top \Xi^\udelta)]^2}{{\bf w}^\top {\bf A}_s^\udelta {\bf w}} \\
    &= \sup_{0 \not= {\bf w} \in \R^{\# Y^\udelta}} \frac{[{\bf w}^\top {\bf r}^\udelta_Y]^2}{{\bf w}^\top {\bf A}_s^\udelta {\bf w}} = \sup_{0 \not= {\bf x} \in \R^{\# Y^\udelta}} \frac{[(({\bf A}_s^\udelta)^{-1/2}{\bf x})^\top {\bf r}^\udelta_Y]^2}{{\bf x}^\top {\bf x}} = \la ({\bf A}_s^\udelta)^{-1} {\bf r}_Y^\udelta, {\bf r}_Y^\udelta \ra. \qedhere
  \end{align*}
\end{proof}
\begin{remark}
  \jw{deze afschatting geldt voro elke $v^\delta \in X^\delta$ en niet alleen voor de oplossing van ons systeem; ik zie twee mogelijk-leuke toepassingen hiervan, namelijk
  \begin{itemize}
    \item het stelt ons in staat om de $X$-norm fout van PCG iterands te plotten
    \item het stelt ons in staat om de $X$-norm fout van de FK19-oplossing te plotten
  \end{itemize}
  is daar iets nuttig aan?}
\end{remark}

@@

\section{Application in linear time}
\jw{XYZ inleiding.} Here
\begin{align*}
(A_s v_1 \otimes w_1)(v_2 \otimes w_2)&=\int_0^1 v_1(t) v_2(t)\dif t \int_\Omega \nabla w_1({\bf x})\cdot  \nabla w_2({\bf x}) \dif{\bf x},\\
(B v_1 \otimes w_1)(v_2 \otimes w_2)&=\int_0^1 \frac{\dif v_1}{\dif t}(t) v_2(t)\dif t \int_\Omega w_1({\bf x}) w_2({\bf x}) \dif {\bf x}+(A_s v_1 \otimes w_1)(v_2 \otimes w_2),\\
(\gamma_0' \gamma_0 v_1 \otimes w_1)(v_2 \otimes w_2)&=
v_1(0) v_2(0)\int_\Omega w_1({\bf x}) w_2({\bf x}) \dif {\bf x}\\
(\gamma_0' u_0)(v_2 \otimes w_2)&=v_2(0) \int_\Omega u_0({\bf x}) w_2({\bf x}) \dif {\bf x}\\
g(v_2 \otimes w_2)&=\sum_i \int_0^1g_{1i}(t) v_2(t)\dif t \int_\Omega g_{2i}({\bf x}) w_2({\bf x}) \dif  {\bf x},
\end{align*}
where we assumed that $g$ is given or approximated in the form $g(t,{\bf x})=\sum_i g_{1i}(t) g_{2i}({\bf x})$.
\subsection{Efficient evaluation of operators in space and time}
TODO
\subsection{Efficient evaluation of tensor-product operators}
TODO
\subsection{Preconditioning the arising system}
\jw{introduce operator preconditioning}

@@

@@

\subsection{Second singular solution}
Selecting the unit-square domain $\Omega := [0,1]^2$ with data $u_0 := \bbone$
and $g \equiv 0$, we see the error progressions plotted in Figure~\ref{fig:singular-error}.
For the uniform refinement, we see the abhorrent convergence rate of $\O(N^{-1/11})$.
Luckily, the adaptive rate is much better at $\O(N^{-1/3})$. It is unknown
if this rate is optimal.

Figure~\ref{fig:singular-time-memory} shows the peak memory consumption in kilobytes
over $\# \dim X^\delta$ and computation time of a single application of $S^{\udelta \delta}$.
We see that the memory consumption is asymptotically linear in the number of
unknowns, whereas for the time per apply, we observe a rate of $1.04$.\jw{dit is natuurlijk groter dan 1. Is het wel acceptabel?}

\begin{figure}
  \includegraphics[width=0.5\linewidth]{singular_adaptive_errors}
  \caption{Error progression of $e^\delta := u - u^{\underline{\delta}\delta}$ for the singular problem as
  a function of $\dim X^\delta$ for adaptive and full grid refinement.\jw{excuse the terrible graph}}
  \label{fig:singular-error}
\end{figure}
\begin{figure}
  \includegraphics[width=\linewidth]{singular_mem_time.pdf}
  \caption{Time and memory progression for the singular problem for adaptive refinement.\jw{excuse the terrible graph}}
  \label{fig:singular-time-memory}
\end{figure}

>>>>

This error estimator will drive our adaptive loop. Even without Assumption~\ref{assum:X-stable-splitting} however, we derive the following efficient and reliable \emph{global estimator}.
\begin{prop}
  \label{prop:global-apost}
  Let $\Xi^{\udelta}$ be a basis for $Y^{\udelta}$. Then, modulo some data-oscillation
  term we expect to be of higher order, for any $v^\delta \in X^\delta$,
  \[
    \|u - v^\delta\|_X^2 \eqsim \la (\mathbf A_s^{\udelta})^{-1}{\bf r}^\udelta_Y, {\bf r}^\udelta_Y \ra + \|u_0 - \gamma_0 v^\delta\|_H^2,
  \] 
  where $\mathbf A_s^{\udelta} := (A_s E_Y^{\udelta} \Xi^{\udelta})(E_Y^{\udelta} \Xi^{\udelta})$ and ${\bf r}^\udelta_Y := (g - B E_X^{\delta} v^\delta)(\Xi^\udelta)$.
  \jw{zouden we in staat zijn om de constantes hidden in $\eqsim$ hierboven, precies te maken? Mss de data oscillation erin?}
\end{prop}
\begin{proof}
  Define $e^\delta := u - E_X^\delta v^\delta$. From~\eqref{eqn:B-lis}, we derive
  \[
    \|e^\delta\|_X^2 \eqsim \|B e^\delta\|_{Y' \times H}^2
    = \|B e^\delta\|_{Y'}^2 + \|\gamma_0 e^\delta\|_H^2 = \|B e^\delta\|_{Y'}^2 + \|u_0 - \gamma_0 v^\delta\|_H^2,
  \]
  and using~\cite[Lemma 3.4]{Stevenson2020a} together with~\eqref{eqn:infinfsup}, we find
  \begin{align*}
    \|B e^\delta\|_{Y'}^2 &= \|e^\delta\|_Y^2 + \|\partial_t e^\delta\|_{Y'}^2 + \|\gamma_T e^\delta\|_H^2 - \|\gamma_0 e^\delta\|_H^2 \\
    &\eqsim \|e^\delta\|_Y^2 + \|\partial_t e^\delta\|_{{Y^\udelta}'}^2 + \|\gamma_T e^\delta\|_H^2 - \|\gamma_0 e^\delta\|_H^2 = \|B e^\delta\|_{{Y^\udelta}'}^2.
  \end{align*}
  The result then follows from
  \begin{align*}
    \|B e^\delta\|_{{Y^\udelta}'}^2 &= \|g - BE_X^\delta v^\delta\|_{{Y^\udelta}'}^2
    = \sup_{\mathclap{0 \not= w \in Y^\udelta}} \frac{[(g - Bv^\delta)(w)]^2}{\|w\|_Y^2} = ~~\sup_{\mathclap{0 \not= {\bf w} \in \R^{\# Y^\udelta}}} ~~\frac{[(g - Bv^\delta)({\bf w}^\top \Xi^\udelta)]^2}{{\bf w}^\top {\bf A}_s^\udelta {\bf w}} \\
    &= \sup_{0 \not= {\bf w} \in \R^{\# Y^\udelta}} \frac{[{\bf w}^\top {\bf r}^\udelta_Y]^2}{{\bf w}^\top {\bf A}_s^\udelta {\bf w}} = \sup_{0 \not= {\bf x} \in \R^{\# Y^\udelta}} \frac{[(({\bf A}_s^\udelta)^{-1/2}{\bf x})^\top {\bf r}^\udelta_Y]^2}{{\bf x}^\top {\bf x}} = \la ({\bf A}_s^\udelta)^{-1} {\bf r}_Y^\udelta, {\bf r}_Y^\udelta \ra. \qedhere
  \end{align*}
\end{proof}
\begin{remark}
  \jw{deze afschatting geldt voro elke $v^\delta \in X^\delta$ en niet alleen voor de oplossing van ons systeem; ik zie twee mogelijk-leuke toepassingen hiervan, namelijk
  \begin{itemize}
    \item het stelt ons in staat om de $X$-norm fout van PCG iterands te plotten
    \item het stelt ons in staat om de $X$-norm fout van (1 component van) de FK19-oplossing te plotten
  \end{itemize}
  is daar iets nuttig aan?}
\end{remark}

@@

\section{Application in linear time}
\jw{XYZ inleiding.}

Here
\begin{align*}
(A_s v_1 \otimes w_1)(v_2 \otimes w_2)&=\int_0^1 v_1(t) v_2(t)\dif t \int_\Omega \nabla w_1({\bf x})\cdot  \nabla w_2({\bf x}) \dif{\bf x},\\
(B v_1 \otimes w_1)(v_2 \otimes w_2)&=\int_0^1 \frac{\dif v_1}{\dif t}(t) v_2(t)\dif t \int_\Omega w_1({\bf x}) w_2({\bf x}) \dif {\bf x}+(A_s v_1 \otimes w_1)(v_2 \otimes w_2),\\
(\gamma_0' \gamma_0 v_1 \otimes w_1)(v_2 \otimes w_2)&=
v_1(0) v_2(0)\int_\Omega w_1({\bf x}) w_2({\bf x}) \dif {\bf x}\\
(\gamma_0' u_0)(v_2 \otimes w_2)&=v_2(0) \int_\Omega u_0({\bf x}) w_2({\bf x}) \dif {\bf x}\\
g(v_2 \otimes w_2)&=\sum_i \int_0^1g_{1i}(t) v_2(t)\dif t \int_\Omega g_{2i}({\bf x}) w_2({\bf x}) \dif  {\bf x},
\end{align*}
where we assumed that $g$ is given or approximated in the form $g(t,{\bf x})=\sum_i g_{1i}(t) g_{2i}({\bf x})$.

\subsection{Efficient evaluation of operators in space and time}
\jw{willen we hier de sectie uit followup copy-pasten of is een referentie naar nicos' phd of dat paper met kestler voldoende?}
\subsection{Efficient evaluation of tensor-product operators}
\jw{willen we hier de sectie uit followup copy-pasten of is een referentie naar nicos' phd of dat paper met kestler voldoende?}

\subsection{Preconditioning the arising system}
\jw{introduce operator preconditioning}

@@

\jw{ik mis nog een paar rake zinnen om te vergelijken met FK19 hieronder. De moeilijkheid
zit 'm in het verschil in norm + dat zelfs hun uniform-refinement solutions tov
een simplicial mesh zijn, en wij dus een tensor mesh hebben}

@@

\subsection{Singular solution}
Selecting the unit-square domain $\Omega := [0,1]^2$ with data $u_0 := \bbone$
and $g \equiv 0$, we see the error progressions plotted in Figure~\ref{fig:singular-error}.
For the uniform refinement, we see the abhorrent convergence rate of $\O(N^{-1/11})$.
Luckily, the adaptive rate is much better at $\O(N^{-1/3})$. It is unknown
if this rate is optimal.

Figure~\ref{fig:singular-time-memory} shows the peak memory consumption in kilobytes
over $\# \dim X^\delta$ and computation time of a single application of $S^{\udelta \delta}$.
We see that the memory consumption is asymptotically linear in the number of
unknowns, whereas for the time per apply, we observe a rate of $1.03$.\jw{dit is natuurlijk groter dan 1. Is het wel acceptabel?}
The memory consumption is relatively high because $\dim Y^\udelta \approx 16 \times \dim X^\delta$,
resulting in very large systems.\jw{oef, ja wel goed om te vertellen maar 16 is een hoge constante zeg}

\begin{figure}
  \includegraphics[width=0.5\linewidth]{singular_adaptive_errors}
  \caption{Error progressions for the singular problem of residual norm (solid line),
  estimated $X$-norm error (dashed), and $t=0$ time-slice error (dotted) as a function
  of $\dim X^\delta$ for adaptive (black), sparse grid (red), and full grid refinement (orange).}
  \label{fig:singular-error}
\end{figure}
\begin{figure}
  \includegraphics[width=\linewidth]{singular_mem_time.pdf}
  \caption{Time and memory progression for the singular problem for adaptive refinement.\jw{excuse the terrible graph}}
  \label{fig:singular-time-memory}
\end{figure}
